### 2014-01-26 Papers_SVN/papers/in_preparation/accepted/2004/coset2004-robert/results/PAPER2.tex

PAPER 2
=======
Reviewer 1
***********
COSET-1 REVIEW FORM
===================

Paper title: "An application oriented parallel programming system for dedicated cluster computing"

Authors: T. Robert C. Santos, A. Augusto Frohlich

Reviewer confidence level (1-4): 3
1=not my area
2=general area knowledge
3=close to my area of expertise
4=my area of expertise

Summary of the paper:

This paper claim in title and sbtract to present a parallel programming system for cluster based on EPOS. But in fact the originality and novelty of this paper concerns the design of lightweight protocols with Myrinet NIC.

Strengths of the paper:

Presents low level protocols with programmable NIC.

Weaknesses of the Paper:
Incomplete introduction adn bibliography. Paper not focused on the real original work.
Miss first experimental results.

Interest for COSET-1:

Paper from south america that could open the world range of COSET workshop.
Paper that could be linked with networking topics.

For the next four queries, please rate the paper on a scale of 1-4, where
the quality increases as the number increases.

Originality:

Technical merit:

Readability:

Overall rating: STRONG ACCEPT (4)  - WEAK ACCEPT (3) - WEAK REJECT (2) - STRONG REJECT (1)
2

Detailed comments for the authors:

The abstract is too general and too vague and should more adress keypoints and challenges present in the paper.

The introduction is also not really focused on the paper and remains vague on the benefits of clusters.
If EPOS is the purpose of the paper, you must cite it in abstract and introduction.

2.2 : "network hardware does not provide..." : you only speak about NIC and not about network equipments (routers, switchs..).

"many network interfaces are equipped..." : not so many : Myrinet is the most faous ! but what about other choices ? you should cite them (Quadrics...)

This paper should be only focused on lightweigth protocols with myrinet NIC and not be interfered by PPS.

You should add first experimental results of your protocols.

Bibliography is incomplete, no conference or journal sources are provided. You must include it in each citation to allow the reader to find cited papers.
***************************************
Reviewer 2
**********

COSET-1 REVIEW FORM
===================

Paper title: An application-oriented parallel-programming system for
             dedicated cluster computing

Authors: Thiago Robert C. Santos, Antonio Augusto Frohlich

Reviewer confidence level (1-4): 3
1=not my area

2=general area knowledge
3=close to my area of expertise
4=my area of expertise

Summary of the paper:

This paper proposes a basic architecture for a family of low-level Myrinet 
protocols. Each protocol in the family is built on top of this architecture 
and is customized to satisfy the needs of a specific application. This
family of protocols is essentially a customizable component in Epos,
authors' previously developed component-based operating system. The idea
behind customizable low-level protocols is that parallel applications running 
on  dedicated clusters can improve their communication performance by using 
the most appropriate protocols for their needs.

Strengths of the paper:

The general idea of customizable OS components is worthwhile.

Weaknesses of the Paper:

Unfortunately, the claims from the abstract and introduction do not match the
actual content of the paper. Most of the paper is a discussion about authors' 
previous work (Epos) and the Myrinet networks in general. It's never clear
what exactly the original contribution is, and why the proposed idea would
perform better than existing approaches. The paper does not present any
implementation details or quantitative evaluation of the proposed
architecture.

Interest for COSET-1:

Although the topic of the paper is of interest for this workshop, the paper 
itself is confusing and lacks originality.

Originality: 1

Technical merit: 1

Readability: 2

Overall rating: WEAK REJECT (2)

Detailed comments for the authors:

The first thing that has to be improved in this paper is its presentation:

- the title does not match the content
- clarify what the original contribution is; in section 3.2, it's never
very clear where the general discussion about previous work ends, and
where your contribution begins.
- if this architecture is supposed to support many low-level Myrinet protocols,
you should present at least one or two protocols to illustrate the idea
(this part is just hand-waved in 3.4).
- how is the architecture different from other architectures based on Myrinet 
networks?
- how about a comparison of Epos with the Pebbles component-based system?

It's very hard to evaluate this paper in terms of technical merit in the 
absence of any implementation and evaluation. Some of the design decisions
seem outrightly wrong. As you mention, the architecture presented in Fig
2 has 4 copy operations for small messages, one more than the minimum
imposed by the Myrinet network. Also, the architecture from Fig 3 should
perform worse than the previous one (one more copy from user to kernel 
and vice-versa for every message).

To make your architecture believable, you definitely need a better description
and motivation of the design decisions as well as a prototype implementation 
to demonstrate its benefits.
************************

Reviewer 3
***********
COSET-1 REVIEW FORM
===================

Paper title: An application-oriented parallel-programming system for
dedicated cluster computing

Authors: Thiago Robert C. Santos and Antonio Augusto Frohlich

Reviewer confidence level (1-4): 4

Summary of the paper:

The paper presents some ongoing work about the communication subsystem
of an application-oriented operating system named EPOS. This OS is
intended to provide the applications with a set of lightweight
communication protocols that can be tailored to match the
applications' needs.

Strengths of the paper:

?

Weaknesses of the Paper:

Clearly fails to present the contribution of the authors: the
description of their environment is not precise enough to get an idea
of what the authors have done. Actually the paper only describes the
Myrinet architecture and the design of existing lightweight protocols
such as BIP...

Interest for COSET-1: Ok.

For the next four queries, please rate the paper on a scale of 1-4, where
the quality increases as the number increases.

Originality: 2

Technical merit: 1

Readability: 4

Overall rating: WEAK REJECT (2)

Detailed comments for the authors:

  The main problem with this paper is that the contribution does not
  clearly appear. It mainly stems from the presentation of the
  authors' implementation which lacks too many details about how the
  lightweight communication protocols can be customized by the
  applications, how the overall software architecture is designed,
  etc. Without these details, it is hard to understand how the EPOS
  system better meet the needs of parallel applications than other
  existing communication subsystems.

  Also, a major weakness of the paper is the lack of performance
  figures and comparison with other systems. Without such an
  evaluation, it is hard to evaluate whether the proposed approach is
  valuable or not.

  Finally, please note that all bibliography entries lack a lot of
  details: name of the conference/journal, volume number, page
  numbers, etc.