### 2014-01-26 Papers_SVN/papers/in_preparation/accepted/2007/euc2007-hybrid_componentes-hugo/latex8.tex

%\documentstyle[times,art10,twocolumn,latex8]{article}

\documentclass[times,10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}

%------------------------------------------------------------------------- 
% take the % away on next line to produce the final camera-ready version 
%\pagestyle{empty}

%------------------------------------------------------------------------- 
\begin{document}

\title{ETFA-2007 Author Guidelines 
       for A4 ({\boldmath $21\times29.7$ cm}) Proceedings Manuscripts}

\author{First Author\\
Institution1\\
First line of institution1 address\\ Second line of institution1 address\\ 
FirstAuthor@institution1.edu\\
% For a paper whose authors are all at the same institution, 
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'', 
% just like the second author.
\and
Second Author\\
Institution2\\
First line of institution2 address\\ Second line of institution2 address\\ 
SecondAuthor@institution2.edu\\
}

\maketitle
\thispagestyle{empty}

\begin{abstract}
The ABSTRACT is to be in fully-justified italicized text, at the top of 
the left-hand column, below the author and affiliation information. Use the word "Abstract" as the title, in 12-point Times, boldface type, 
centered relative to the column, initially capitalized. The abstract is to 
be in 10-point, single-spaced type. The text of the abstract should not be 
more than 16 lines long (this abstract is exactly sixteen lines long). 
Abstracts should be followed by one blank line (using the abstract style) 
if the text begins with a level 1 heading (like here) or two blank lines 
if the text begins with the normal style.

These last five lines are just to fill out the abstract text to exactly 16 
lines. These last five lines are just to fill out the abstract text to 
exactly 16 lines. These last five lines are just to fill out the abstract 
text...
\end{abstract}

%------------------------------------------------------------------------- 
\Section{Introduction}

Please follow the steps outlined below when submitting your final 
manuscript to the ETFA conference.

%------------------------------------------------------------------------- 
\Section{Instructions}

Please read the following carefully.

%------------------------------------------------------------------------- 
\SubSection{Language}

All manuscripts must be in English. The use of an automatic spell-checker 
is advised.

%------------------------------------------------------------------------- 
\SubSection{Submitting your paper}

The final manuscript of your paper must be submitted through the electronic 
submission systems on the conference web page, in pdf format. Please 
verify if it prints properly on A4 printer paper.

%------------------------------------------------------------------------- 
\SubSection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must be 
kept within a print area 17 cm wide by 24.2 cm high. Do not write or print 
anything outside the print area. Do not include any page numbering in your 
manuscript.

%------------------------------------------------------------------------ 
\SubSection{Length of your paper}

The length of your paper is limited to 8 two-column pages. For any extra 
proceeding pages please check the conference web page.

%------------------------------------------------------------------------ 
\SubSection{Formatting your paper}

All text must be in a two-column format. The total allowable width of the 
text area is 17 cm wide by 24.2 cm high. Columns are to be 8.1 cm wide, 
with 0.8-cm space between them. The main title (on the first page) should 
begin 3 cm from the top edge of the page. The second and following pages 
should begin 2.5 cm from the top edge. On all pages, the bottom margin 
should be, at least, 3 cm from the bottom edge of the page.

%------------------------------------------------------------------------- 
\SubSection{Type-style and fonts}

Wherever Times is specified, Times Roman may also be used. If neither is 
available on your word processor, please use the font closest in 
appearance to Times that you have access to.

MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of 
the first page. The title should be in Times 14-point, boldface type. 
Capitalize the first letter of nouns, pronouns, verbs, adjectives, and 
adverbs; do not capitalize articles, coordinate conjunctions, or 
prepositions (unless the title begins with such a word). Leave two blank 
lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title 
and printed in Times 12-point, non-boldface type. This information is to 
be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use 
double-spacing. All paragraphs should be indented 1 pica (approx. 1/6 
inch or 0.422 cm). Make sure your text is fully justified---that is, 
flush left and flush right. Please do not place any additional blank 
lines between paragraphs. Figure and table captions should be 10-point 
Helvetica boldface type as in
\begin{figure}[h]
   \caption{Example of caption.}
\end{figure}

\noindent Long captions should be set as in 
\begin{figure}[h] 
   \caption{Example of long caption. It is 
     aligned on both sides and indented with an 
     additional margin on both sides of 0.42-cm (1~pica).}
\end{figure}

\noindent Initially capitalize only the first word of section titles and 
first-, second-, and third-order headings.

FIRST-ORDER HEADINGS should be Times 12-point boldface, initially 
capitalized, flush left, with one blank line before, and one blank line 
after.

SECOND-ORDER HEADINGS should be Times 10-point boldface, initially 
capitalized, flush left, with one blank line before, and one after. If you 
require a third-order heading (we discourage it), use 10-point Times, 
boldface, initially capitalized, flush left, preceded by one blank line, 
followed by a period and your text on the same line.

%------------------------------------------------------------------------- 
\SubSection{Footnotes}

Please use footnotes sparingly%
\footnote
   {%
     Or, better still, try to avoid footnotes altogether.  To help your 
     readers, avoid using footnotes altogether and include necessary 
     peripheral observations in the text (within parentheses, if you 
     prefer, as in this sentence).
   }
and place them at the bottom of the column on the page on which they are 
referenced. Use Times 8-point type, single-spaced.

%------------------------------------------------------------------------- 
\SubSection{Illustrations and graphs}

All illustrations and graphs should be centered and must be in place in 
the manuscript. If you must use photographs, they must be scanned and 
pasted onto your manuscript. Please supply the best quality photographs 
and illustrations possible.

%------------------------------------------------------------------------- 
\SubSection{References}

List and number all bibliographical references in 9-point Times, 
single-spaced, at the end of your paper. When referenced in the text, 
enclose the citation number in square brackets, for example~\cite{ex1}. 
Where appropriate, include the name(s) of editors of referenced books.

%------------------------------------------------------------------------- 
\nocite{ex1,ex2}
\bibliographystyle{latex8}
\bibliography{latex8}

\end{document}