### 2014-01-26 Papers_SVN/papers/in_preparation/accepted/2009/tecs2009-pm-guto/conclusion.tex

\section{Conclusion}\label{sec:conclusion}

In this article, power management in embedded systems was addressed from
energy-aware design to energy-efficient implementation, aiming at
introducing a set of mechanisms specifically conceived for this
scenario. A power management API defined at the level of user-visible
system components was proposed and compared with traditional APIs. Its
implementation was discussed in the context of the necessary
infrastructure, including battery monitoring, accounting, auto-suspend,
and auto-resume. An energy-event propagation mechanism based on
\emph{Petry Nets} was proposed and its implementation using
\emph{Aspect-Oriented Programming} techniques was depicted. The use of
the proposed infrastructure by an autonomous power manager integrated
into a real-time scheduler was also discussed, thus covering the main
components of a modern power management system for embedded systems.

% The proposed strategy was illustrated and evaluated through a didactic,
% yet realistic, example embedded system targeted at environment
% temperature monitoring. The example was described from early design
% stages down to a real implementation for the \textsc{Epos} system on the
% Mica2 Mote, a well-known platform that helps to put the proposal into
% perspective.  Experiments with this implementation showed that power
% management is an expensive service for simple embedded systems,
% specially when provided in the context of hard real-time applications
% and with limited hardware support.

The proposed strategy was illustrated and evaluated through a didactic,
yet realistic, example embedded system targeted at environment
temperature monitoring. The example was described from early design
stages down to a real implementation for the EPOS system on the Mica2
Mote, a well-known platform that helps to put the proposal into
perspective. Experiments with this implementation showed that
integrating the proposed power management mechanisms into a hard-real
time run-time support system comes at a high cost in terms of program
memory, specially in platforms with limited hardware support.
Nonetheless, they also showed a relatively small impact on performance,
slightly adding to the latency of scheduling and I/O operations. This
can be explained mainly by the sporadic nature of power management
operations.

The experiments carried out also made evident the benefits of the
proposed mechanisms in terms of energy efficiency and system utility as
they confirmed the strategy's ability to sustain a given lifetime for
the system without affecting the deadlines of hard real-time tasks at
the same time it enables the safe usage of the remaining energy by
best-effort tasks. These benefits arise from proposed strategy itself
and are not dependent from \textsc{Epos} or Mica2. Therefore, the
intended contribution for this article is not ``yet another power
manager for embedded systems'', but the introduction of a broader and
systematic way to deal with power management issues in embedded systems.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "pm"
%%% End: