### 2014-01-26 Papers_SVN/papers/in_preparation/accepted/2010/ees-timer-guto/coverLetter.tex

\documentclass[english,11pt]{letter}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{babel}
\usepackage[top=.5in,bottom=.5in,left=1in,right=1in,noheadfoot]{geometry}
\usepackage{graphicx}

\address{\sffamily
  \begin{tabular}{rr}
    Prof.~Dr.~Antônio Augusto Fröhlich & \raisebox{-0.9cm}[0pt][0pt]{\resizebox*{!}{1.2cm}{\includegraphics{/home/guto/etc/logos/lisha}}}\\
    Federal University of Santa Catarina &\\
    Software/Hardware Integration Laboratory &\\
    \multicolumn{2}{r}{PO Box 476, 88040-900 Florianópolis, SC, Brazil}\\
    \multicolumn{2}{r}{Phone: +55 48 3721-9516}\\
    \multicolumn{2}{r}{E-Mail: guto@lisha.ufsc.br}\\
    \multicolumn{2}{r}{WWW: http://lisha.ufsc.br/$\sim$guto}\\
    \\
    \multicolumn{2}{r}{\today}
  \end{tabular}
}

\date{}

\signature{
  \begin{center}
   \raisebox{-.9cm}[0pt][0pt]{\includegraphics[scale=1]{/home/guto/personal/signature}}\\
    \hrulefill\\
    Antônio Augusto Fröhlich
  \end{center}
}

\begin{document}

\begin{letter}{ELSEVIER Computers & Electrical Engineering\\
    Editor-in-Chief\\
    Professor Dr. Manu Malek  }
\opening{Dear Prof. Malek,}
\setlength{\parskip}{1.5ex}

I would like to submit the attached manuscript, "Periodic Timers
Revisited: the Real-time Embedded System Perspective", for consideration
for possible publication in the ELSEVIER Computers & Electrical Engineering.

Regarding the option of nominating reviewers, I would like
to suggest two nominees:

\begin{tabular}{p{.45\textwidth} p{.45\textwidth}}
  Hermann Kopetz & Franz Rammig\\
Technische Universtität Wien & Universität Paderborn\\
Institut für Technische Informatik & Heinz Hixdorf Institute\\
+43 1 58801 / 18210 & +49 5251 606500\\
hk@vmars.tuwien.ac.at & franz@upb.de\\
\end{tabular}

Both Prof. Kopetz and Prog. Rammig have a long publication list on
real-time embedded systems and have advised PhDs in areas close to
that of the manuscript. The nominees for editor and reviewers have no
conflict of interest with any of the authors of the paper being
submitted.

This paper has not been published or accepted for publication and it
is not under consideration at another journal. Although we mention a
previous paper on the subject in the manuscript, we do not consider
this to be an extension of that. The first was mostly a motivation for
the experiments described in the manuscript.  Nevertheless, should it
be consider an extension by the editorial board, it is certainly far
above the 25\% new content line.

Sincerely,

\closing{~}

\end{letter}

\end{document}