### 2014-01-26 Papers_SVN/papers/in_preparation/accepted/2011/smc2011-energy_accounting-arliones/files/6_conclusion.tex

\section{Conclusion}
\label{sec:concl}

We presented a software implementation of an energy consumption accounter for
battery-operated embedded systems. We modeled the accounter and implemented it
in a simulation environment and in a real platform
(\epos~\cite{Project:EPOS:2010}), altought the implementation details in the
real platform have been kept out due to the page limit. In order to lower the
processing overhead imposed by our approach we extracted runtime parameters of
energy consumption and execution time of a given application and submitted it to
a meta-heuristic optimizer (\nsga). The objective is to look for good solutions
for the period at which battery-related information should be updated in order
to maximize system utilization while minimizing residual energy and guaranteeing
a pre-defined system lifetime (mission duration). A case study on an IP-based
network running over IEEE 802.15.4 sensing nodes showed promising results. It
showed that the approach was able to control energy consumption on the network
in a way that none of the nodes ran out of battery before the pre-defined
mission duration has elapsed.

On going studies are extending the work in this paper by deploying fairer
scheduling mechanisms to reduce the impact on system quality. This effort rely
on flexible task scheduling schemes
% , such as \textit{(m,k)-firm}~\cite{Ramanathan:1997} or the elastic
% model~\cite{Buttazzo:1998},
to be put in place of the current egoist approach of preventing tasks'
execution and using network-wide battery charge information provided by the
accounter described herein as parameter for \adhop's pheromone generation
function.