### 2014-01-26 Papers_SVN/papers/rejected/2007/2007-pm-arliones-eurosys/power.tex

\documentclass[preprint]{sigplanconf}

\usepackage[T1]{fontenc}
\usepackage[english]{babel}   
\usepackage[latin1]{inputenc}  
\usepackage{times}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{cite}

\lstloadlanguages{[ANSI]C++,HTML}
\lstdefinelanguage{XML} {
  keywords={xml,version,DOCTYPE,SYSTEM,EPOSConfig,family,member,name,type,
  feature,trait,class,performance,codesize,energy,power,default,pos,pre}}
\lstdefinestyle{prg} {basicstyle=\small\sffamily, lineskip=-0.2ex}
\lstdefinestyle{prgbox} {basicstyle=\small\sffamily lineskip=-0.2ex}
\lstdefinestyle{inlineprg} {basicstyle=\small\sffamily}

\newcommand{\fig}[4][htbp]{
  \begin{figure}[#1] {\centering{\scalebox{#4}{\includegraphics{fig/#2}}}\par}
    \caption{#3\label{fig:#2}}
  \end{figure}
}

\newcommand{\largefig}[4][htb]{
  \begin{figure*}[#1] {\centering{\scalebox{#4}{\includegraphics{fig/#2}}}\par}
    \caption{#3\label{fig:#2}}
  \end{figure*}
}

\newcommand{\prg}[4][htb]{
  \begin{figure}[#1]
%    \vspace{\parskip}
%    \makebox[\textwidth][c]{
      \lstinputlisting[language=#2,style=prg]{prg/#3.prg} %}
%    \vspace{0.4\parskip}
    \caption{#4\label{prg:#3}}
  \end{figure}
}

%\newcommand{\note}[1]{\marginpar{\footnotesize{#1}}}
\newcommand{\note}[1]{}

\begin{document}

\conferenceinfo{EuroSys '07}{Lisbon}
\copyrightyear{2005}
\copyrightdata{[to be supplied]}

\title{Application-Driven Power Management for Embedded Systems}

\authorinfo{Arliones Stevert Hoeller Junior, Lucas Francisco Wanner and Antnio Augusto Frhlich}
           {Laboratory for Software and Hardware Integration\\P.O.Box 476, 88040900\\Florianpolis - Brazil}
           {\{arliones,lucas,guto\}@lisha.ufsc.br}

\maketitle

\begin{abstract}

  Deeply Embedded Systems usually are simple, battery-powered systems
  with resource limitations. In some situations, their batteries
  lifetime becomes a primordial factor for reliability. Because of
  this, it is very important to handle power consumption of such
  devices in a non-restrictive and low-overhead way. This power
  management cannot restrict the wide variety of different low-power
  modes such devices often feature, thus allowing a wider system
  con\-fi\-gu\-ra\-bi\-li\-ty. However, once in such devices
  processing and memory are often scarce, the power management
  strategy cannot compromise large amounts of system resources. In
  this paper we propose a simplified interface for power management
  of software and hardware components. The approach is based on the
  hierarchical organization of such components in a component-based
  operating system and allows power management of system components
  without the need for costly techniques or strategies. A case study
  including real implementations of system and application is
  presented to evaluate the technique and shows energy saves of
  almost 40\% by just allowing applications to express when certain
  components are not being used.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introdução}
\label{sec:introducao}
\input{intro}

\section{O gerente de consumo de energia proposto}
\label{sec:gerente}
\input{gerente}

\section{Implementation}
\label{sec:implementacao}
\input{implementacao}

\section{Case study}
\label{sec:estudosdecaso}
\input{estudosdecaso}

\section{Related Work}
\label{sc:related}
\input{related.tex}

\section{Conclusion}
\label{sec:conclusao}
\input{conclusion}

%\appendix

%\section{Rede de migrações de modos de operação gereralizada}
%\label{apx:general_net}
%\begin{figure} {\centering{\scalebox{0.55}{\rotatebox{270}{\includegraphics{fig/general_net}}}}\par}
%\end{figure}

\acks

Authors would like to thank Augusto Born de Oliveira, Hugo Marcondes,
Rafael Cancian and, specially, Prof. Jlio Zseremetta from Federal
University of Santa Catarina for very helpful discussion.

\bibliographystyle{plain}
\bibliography{power}

\end{document}