### 2014-01-26 Papers_SVN/papers/rejected/2011/2011-hpcs-giovani-hipeac/coverLetter/coverLetter.tex

\documentclass[english,11pt]{letter}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{babel}
\usepackage[top=.5in,bottom=.5in,left=1in,right=1in,noheadfoot]{geometry}
\usepackage{graphicx}

\address{\sffamily
  \begin{tabular}{rr}
    Prof.~Dr.~Antônio Augusto Fröhlich & \raisebox{-0.9cm}[0pt][0pt]{\resizebox*{!}{1.2cm}{\includegraphics{lisha}}}\\
    Federal University of Santa Catarina &\\
    Software/Hardware Integration Laboratory &\\
    \multicolumn{2}{r}{PO Box 476, 88040-900 Florianópolis, SC, Brazil}\\
    \multicolumn{2}{r}{Phone: +55 48 3721-9516}\\
    \multicolumn{2}{r}{E-Mail: guto@lisha.ufsc.br}\\
    \multicolumn{2}{r}{WWW: http://lisha.ufsc.br/$\sim$guto}\\
    \\
    \multicolumn{2}{r}{\today}
  \end{tabular}
}

\date{}

\signature{
  \begin{center}
%   \raisebox{-.9cm}[0pt][0pt]{\includegraphics[scale=1]{/home/guto/personal/signature}}\\
    \hrulefill\\
    Antônio Augusto Fröhlich
  \end{center}
}

\begin{document}

\begin{letter}{ACM Transactions on Architecture and Code Optimization (TACO)\\
    Editor-in-Chief\\
    Professor Dr. Tom Conte  }
\opening{Dear Prof. Conte,}
\setlength{\parskip}{1.5ex}

I would like to submit the attached manuscript, ``Monitoring Shared Memory Invalidations in Embedded Multicore Software through HPCs'', for consideration for possible publication in the ACM Transactions on Architecture and Code Optimization through the 7th International Conference on High-Performance and Embedded Architectures and Compilers (HiPEAC 2012).

This paper has not been published or accepted for publication and it
is not under consideration at another conference or journal.

Regarding the list of conflicts, the two authors of this paper do not have any association with the TACO Associate Editors. The authors' affiliation and their collaboration in the last 5 years are the following:

\begin{tabular}{p{.45\textwidth} p{.45\textwidth}}
  \textbf{Prof. Dr. Ant\^{o}nio Augusto Fr\"{o}hlich} & \textbf{M.Sc. Giovani Gracioli}\\
Ph.D in Computer Engineering from the Technical University of Berlin, Germany 2001 (Advised by Prof. Dr. Wolfgang Schr\"{o}der-Preikschat) & Bachelor degree in Computer Science from the Federal University of Santa Maria, Brazil 2007 (Advised by Prof. Dr. Raul Cereta Nunes)\\

Current associate professor at Federal University of Santa Catarina, Brazil & M.Sc. degree in Computer Science from the Federal University of Santa Catarina, Brazil 2009 (Advised by Prof. Dr. Ant\^{o}nio Augusto Fr\"{o}hlich)\\

Visitor Professor at University of Paderborn, Germany 2007 (research group from Prof. Dr. rer. nat. Franz Josef Rammig) & Current Ph.D student in Control and Automation Engineering at Federal University of Santa Catarina, Brazil (Advised by Prof. Dr. Ant\^{o}nio Augusto Fr\"{o}hlich)\\

\multicolumn{2}{p{.9\textwidth}}{Cooperation project with the Real-time Embedded Software Group (headed by Prof. Dr. Sebastian Fischmeister) from University of Waterloo, Canada, 2011.}

\end{tabular}

Sincerely,

\closing{~}

\end{letter}

\end{document}