### 2014-01-26 Papers_SVN/papers/rejected/2011/2011-security-rodrigo-ipccc/paper.tex

\documentclass[conference]{IEEEtran}

\usepackage[utf8]{inputenc}	% for Latin languages
\usepackage[T1]{fontenc}	% for ISO and UTF characters
\usepackage[english]{babel}	% for multilingual support
\usepackage{graphicx}
\usepackage{subfig}

\newcommand{\fig}[4][htbp]{
  \begin{figure}[#1] {\centering\scalebox{#2}{\includegraphics{fig/#3}}\par}
    \caption{#4\label{#3}}
  \end{figure}
}

\begin{document}

\title{A Trustful Infrastructure for Wireless Sensor Networks}

\author{
    \IEEEauthorblockN{Rodrigo Vieira Steiner, Leonardo Maccari Rufino, and Antônio Augusto Fröhlich}\\
    \IEEEauthorblockA{
  Software/Hardware Integration Lab\\
  Federal University of Santa Catarina\\
  PO Box 476, 88040-900 - Florianópolis, SC, Brazil \\
            \{rodrigo,leonardo,guto\}@lisha.ufsc.br
    }
}

\maketitle

\begin{abstract}
This paper describes the design, implementation and evaluation of a trustful infrastructure for Wireless Sensor Networks (WSNs).
This infrastructure is composed by an embedded platform, EPOSMote II, and a communication protocol stack (TCP/IP/C-MAC).
We focus on security and reliability while still compromised with the low utilization of resources available in sensor nodes. 
%To achieve confidentiality, integrity, and authentication, we make use of data encryption and a temporal authentication scheme.
To achieve confidentiality, integrity, and authentication, we make use of data encryption and authentication.
Differently from other approaches our authentication scheme possesses a temporal property which protects the network against replay attacks.
Our proposal is evaluated in terms of processing time, memory footprint, and energy consumption presenting very little overhead.

%We make use of the hardware-assisted security mechanism present in EPOSMote II to encrypt and decrypt all transmitted data using the Advanced Encryption Standard (AES).
%Although communication in WSNs is error-prone, TCP provides reliable and ordered packet delivery.
%As to not redo TCP's work we make use of the Configurable Medium Access Control (C-MAC) in the data link layer.
%To this end we configured C-MAC in a very simplistic form, using only CSMA and backoff periods.
%Furthermore, we make use of the Hardware-assisted security mechanism present in EPOSMote II to encrypt and decrypt all transmitted data using the Advanced Encryption Standard (AES).
\end{abstract}

\begin{IEEEkeywords}
  security; WSN; authentication;
\end{IEEEkeywords}

\input{files/introduction}

%\input{files/section_ii}

\input{files/proposal}

\input{files/results}

\input{files/conclusions}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}