### 2014-01-26 Papers_SVN/papers/rejected/2011/2011-security-rodrigo-iwcmc/paper.tex

\documentclass[conference]{IEEEtran}

\usepackage[utf8]{inputenc}	% for Latin languages
\usepackage[T1]{fontenc}	% for ISO and UTF characters
\usepackage[english]{babel}	% for multilingual support
\usepackage{graphicx}
\usepackage{subfig}

\newcommand{\fig}[4][htbp]{
  \begin{figure}[#1] {\centering\scalebox{#2}{\includegraphics{fig/#3}}\par}
    \caption{#4\label{#3}}
  \end{figure}
}

\begin{document}

\title{On the Way to a Trustful Embedded Infrastructure for Future Internet}

\author{
    \IEEEauthorblockN{Rodrigo Vieira Steiner, Alexandre Massayuki Okazaki, Alex de Magalhães Machado, \\Giovani Gracioli, and Antônio Augusto Fröhlich}\\
    \IEEEauthorblockA{
  Software/Hardware Integration Lab\\
  Federal University of Santa Catarina\\
  PO Box 476, 88040-900 - Florianópolis, SC, Brazil \\
            \{rodrigo,alexandre,alex,giovani,guto\}@lisha.ufsc.br
    }
}

\maketitle

%Summary: FP7 - IoT
%Conferência alvo: IWCMC2011
%Deadline: 15/12/2010

%Nature: Trustworthiness on Future Internet
%Scope: Named Data Networking executing on EPOSMote II
%Motivation: It is a mistake to assume that each embedded object of Future Internet will implement the TCP/IP stack, especially considering aspects such as power consumption, memory requirement, size, and cost. Furthermore, the IP protocol itself, the driving force of technological revolution which we live, seems to have reached its limit when it comes to Internet and the Future Internet of Things.
%Contribution: This work proposes a Trustful Infrastructure for Future Internet, consisting of an embedded platform (EPOSMote II) and a stack of communication protocols (UDP@NDN@C-MAC) focusing on trustworthiness.

%Title: On the Way for a Trustful Infrastructure for Future Internet

\begin{abstract}
%+ Motivation
%todo
It is a mistake to assume that each embedded object of Future Internet will implement the TCP/IP stack, especially considering aspects such as power consumption, memory requirement, size, and cost.
Furthermore, the IP protocol itself, the driving force of technological revolution which we live, seems to have reached its limit when it comes to the Future Internet.
%+ Problem
%todo
One of the challenges in designing Future Internet applications, however, is the very distinct basic software (e.g OS, communication protocols) that may be present in different devices, and how to combine them in order to build a coherent final application.
The current research and development are focusing on Internet based communication protocols, and on top of that, new protocols are defined to address the target service.
Safety and security capabilities in this domain are open issues that need to be addressed.
Nevertheless, implementing traditional secure communication approaches in resource-constrained platforms, such as those found in Future Internet, is not a straightforward task.
%+ Contribution
In this article, we discuss the limitations of the TCP/IP stack and present an alternative that might be used for Future Internet.
Our solution is composed by an embedded platform, EPOSMote II, and a stack of communication protocols (UDP@NDN@C-MAC) designed specifically to guarantee a trustful communication while still compromised with the low utilization of resources.
\end{abstract}

\input{files/introduction}

\input{files/section_ii}

\input{files/proposal}

\input{files/conclusions}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}