### 2014-01-26 Papers_SVN/papers/rejected/2012/ase2012-test_automation-rita/paper.tex

\documentclass{acm_proc_article-sp}

\usepackage[utf8]{inputenc}	% for Latin languages
\usepackage[T1]{fontenc}	% for ISO and UTF characters
\usepackage[english]{babel}	% for multilingual support

\newcommand{\fig}[4][thb]{
  \begin{figure}[#1] {\centering{\includegraphics[#4]{fig/#2}}\par}
    \caption{#3\label{fig:#2}}
  \end{figure}
}

\begin{document}

\title{Using GDB and QEMU for cross debugging through automatic exchange of configuration parameters.}

\numberofauthors{2}
\author{
\alignauthor
Rita de Cássia Cazu Soldi\\
       \affaddr{Federal University of Santa Catarina(UFSC)}\\
       \affaddr{Laboratory for Software and Hardware Integration(LISHA)}\\
       \affaddr{P.O. Box 476}\\
       \affaddr{88049-900 - Florianópolis, SC, Brazil}\\
       \email{rita@lisha.ufsc.br}
% 2nd. author
\alignauthor
Antônio Augusto Medeiros Fröhlich\\
       \affaddr{Federal University of Santa Catarina(UFSC)}\\
       \affaddr{Laboratory for Software and Hardware Integration(LISHA)}\\
       \affaddr{P.O. Box 476}\\
       \affaddr{88049-900 - Florianópolis, SC, Brazil}\\
       \email{guto@lisha.ufsc.br}
}

\maketitle
\begin{abstract}
   % Theme:  Automated test
   % Scope:  Automated execution tests for helping developer debug the source code.
   % Goal:   Automated test execution by modifying parameters of application and reporting to the developer.  
Debugging is one of the most time-consuming process in software developmen. The debugging task for embedded systems brings some additional problems in the process, since the developer should perform it in different platforms, which vary accordingly to the architecture and vendor. 
This paper presents the automatic exchange of configuration parameters as one anatomic part of fully automated debug. Also, we construct an environment for cross debugging embedded applications based on specific hardware/software requirements.
A real-world application was debugged using the proposed automation in a tailored environment. The obtained results indicates that even a small part of the fully automated debug produce answers to help developers find and fix a bug.
\end{abstract}

%\category{D.2.5}{SoftwareEngineering}{Testing and Debugging}[Debugging Aids, Testing tools]
%\category{D.2.9}{Software Engineering}{Management}[Life cycle, Software quality assurance]
%\category{D.4.5}{Operating Systems}{Metrics}[Reliability]

%\terms{Reliability, Verification}

\keywords{Test automation, cross debug, embedded systems,}

\input{files/introduction}
\input{files/related_work}
\input{files/environment}
\input{files/case_study}
\input{files/evaluation}
\input{files/conclusion}

\bibliographystyle{abbrv}
\bibliography{references}

\balancecolumns
\end{document}