### 2014-01-26 Papers_SVN/papers/rejected/2012/dac2012-unified_design-tiago/tab/tab_sw_metrics_reduced.tex

\begin{tabular}{|l|ccc|rrr|}
\cline{2-7}
\multicolumn{1}{c}{}  &   \multicolumn{3}{|c|}{Memory footprint (code/data) in bytes}    & \multicolumn{3}{|c|}{Avg. operation exec. time in $\mu$s} \\
\multicolumn{1}{c|}{} &   SW-only    &   Unified     &       Diff. (\%)                  & SW-only   &  Unified  &     Diff. (\%) \\
\cline{1-1}
Scheduler             &  8980 / ~432 &   9184 / ~432 &  02.27 / ~00.00                    & 4.26      &   4.52    & 06.10\\
ADPCM                 &  1992 / ~~72 &   2220 / ~~56 &  11.44 / -22.22                   & 3.90      &   3.90    & 00.00\\
DTMF                  &  1076 / 3152 &   1536 / 2884 &  42.75 / -08.50                   & 6844.90   &   7484.00 & 09.35\\
\hline

\end{tabular}

%SW AVGs
%Memory footprint = 4.19 %
%Exec. time = 5.15 %