### 2014-01-26 Papers_SVN/papers/rejected/2012/date2013-dmec-mateus/tex/cell.tex

% ------------------------------------------------------------------------------
\subsection{Cell BE}
% ++ Cell BE
% Describes how data transference is implemented: DMA between PPE and SPEs.
% Describes how synchronization is implemented: Barriers like mechanisms using
% Cell MailBox

% = What are our goals in evaluating this hardware implementation.
The Cell BE is a processor architecture developed by IBM,
Sony, and Toshiba, targeting applications with high thread-level parallelism \cite{Gschwind:2006}.
Our goal on implementing DMEC at Cell BE is evaluating ME in a high performance
and distributed architecture.

% = Target architecture
Cell BE is an example of a
``Multi-computer-on-a-Chip'' architecture because it is composed by distinct
processor units, each one having its own memory address space.
The Cell BE architecture incorporates in the same chip nine independent cores:
a Power Processor Element (PPE) and eight
Synergistic Processor Elements (SPEs).
% 
The PPE is the main processing unit, responsible for managing all chip resources.
The SPEs are dedicated processing units supporting vectorized
floating point code execution.
Each SPE has its own local memory, called Local Storage (LC).
Hence, all communication between cores is performed explicitly though the
Element Interconnect BUS (EIB), a high performance bus.

% \fig{.4}{cell}{Cell BE architecture}

% = Memory model / Data transference
As Cell BE uses memory blocks in distinct address spaces,
we have mapped the operations of \emph{TransferenceManager} to
DMA requests using the EIB.
Using DMA requests, each \emph{Worker} module of DMEC obtain the samples
it needs to perform the ME and deliver back the computed ME results 
to the \emph{Coordinator} module.
% 
The local memory of SPEs is limited to 256KB, including data and code.
This imposed several limitations. For instance, using a partition mode with
six partitions of 640x544 each at a 1080p resolution requires 680KB for data storage.
Thus, only the samples that are currently been used
by ME computation (or samples that are going to be used in a near future) are
kept inside the SPEs local memory.
The samples transferences are hidden from the ME algorithm by the implementation of
the \emph{Picture} interface. The \emph{Picture}, in this case, has a local buffer
used as a cache and obtains sample as needed through the \emph{TransferenceManager}.
% Avaliamos mecanismos de \emph{double buffer} afim de esconder a latência das 
% transferências de memória via DMA entre a memória principal e a memória local 
% das SPEs.

Similarly, the implementation of \emph{PictureMotionCounterpart} for Cell BE
also uses DMA requests through \emph{TransferenceManager}, in order to
deliver the ME results to the main memory.
% TODO maybe move this to DMEC Section.
During the ME computation, previously calculated motion vectors can be used to
improve the ME for the current blocks.
Such motion vectors are called \emph{motion predictors}.
The implementation of \emph{PictureMotionCounterpart} for Cell BE contains an
internal buffer to keep motion predictors inside the local memory of
the \emph{Worker} module, avoiding unnecessary DMA transferences between
the \emph{Worker} and the \emph{Coordinator} modules.

% = Synchronization
The \emph{SynchronizationManager} interface, which specifies barrier-like
mechanisms, is implemented in Cell BE using the \emph{MailBox}, a
hardware resource present on each SPE.
By using specific MailBox messages, it is possible to specify when the
\emph{Worker} has completed the ME computation as well when there are new 
pictures to be processed.

% ------------------------------------------------------------------------------