### 2014-01-26 Papers_SVN/papers/rejected/2012/iccad2012-unified_design-tiago/tab/tab_sw_metrics_reduced.tex

\begin{tabular}{|l|ccc|rrr|}
\cline{2-7}
\multicolumn{1}{c}{}  &   \multicolumn{3}{|c|}{Memory footprint (code/data) in bytes}    & \multicolumn{3}{|c|}{Avg. operation exec. time in $\mu$s} \\
\multicolumn{1}{c|}{} &   SW-only    &   Unified     &       Diff. (\%)                  & SW-only   &  Unified   &     Diff. (\%) \\
\cline{1-1}
Scheduler             &  8980 / ~432 &   9184 / ~432 &  ~2.27 / ~~0.00                   & ~~~4.26   &   ~~~~4.52 & ~6.10\\
ADPCM                 &  ~738 / ~436 &   ~874 / ~436 &  15.56 / ~~0.00                   & ~~~3.90   &   ~~~~3.90 & ~0.00\\
DTMF                  &  ~698 / 3152 &   ~726 / 3152 &  ~3.85 / ~~0.00                   & 6844.90   &   7484.00  & ~9.35\\
\hline

\end{tabular}

%SW AVGs
%Memory footprint = 4.19 %
%Exec. time = 5.15 %