### 2014-01-26 Papers_SVN/papers/rejected/2012/scp2012-mpl_binding_generation-mateus/tex/relat.tex

% + Related Work
%
% This section reviews how hardware devices can be abstracted and organized:
% it reviews the concept of HAL and it is shown why HAL (monolith) is not
% suitable for embedded systems.
% Then, hardware mediators are presented as an alternative to HAL.
%
% Then it presents the concept of FFI (how MPLs interact with hardware devices)
% and show its limitations: it does not guide the developers on the abstraction
% of hardware devices, manual utilization of an FFI can be error
% prone (manual parsing), developer must take into consideration the semantics
% of two distinct languages.
% 
% Then it presents the approaches to facilitate such interaction.
% Generators are presented as a alternative, but they not focus on hardware
% abstraction.
%
% [Talvez apresentar as outras coisas que existem além de FFI... ver aquele
% paper "desmistifing magic" ...]
% [Frameworks maybe
% such as KESO FFI ...(no... I'm going to considerer just as a FFI) ]
%
% HAL Java is another approach...(hardware objects)
%
%
% Resumindo:
% \ref{sec:relat} reviews how hardware devices can be abstracted and organized,
% how MPLs interact with hardware devices, and what are the approaches to
% facilitate such interaction.
% 
% ------------------------------------------------------------------------------
\section{Related work} \label{sec:relat}
% In order to understand how hardware devices can be used
% 
% Two research fields are direct related to our work: how to abstract and organize
% hardware devices to be used by high-level languages which direct memory access
% support, such as C and C++,  and how these abstractions relate to MPLs.
%
Two research fields are direct related to our work: how to abstract and organize
hardware devices to be used by software, and how make these abstractions
accessible to MPLs.
% 
% This section reviews how MPL interact to hardware devices, and how hardware
% devices are abstracted and organized to be used by software.

\subsection{Hardware abstraction}
\emph{Hardware Abstraction Layer} (HAL) is an approach to abstract
hardware specificities, providing the operating system (OS)
with abstract hardware devices.
The hardware device abstraction is performed by services such as interruption
handling, re-initialization handling, DMA transferences, timers' control, and
multiprocessors synchronization \cite{Tanenbaum:2007:MOS:1410217}.

In order to port a HAL to a new hardware platform,
it is necessary to implement all services provided by such HAL.
However, many services provided by a HAL might not be used by an application
and providing all these services generates unnecessary memory and performance
overhead.
% 
An example of overhead caused by the way HALs are designed and implemented is
the \textsc{uCLinux}, an OS which target embedded systems. % CITE uCLinux.
\textsc{uCLinux} inherits several of its features from \textsc{Linux},
among them, the file system abstraction.
Such fact impacts not only on the OS size, but also in all system
initialization infrastructure and application loading \cite{Polpeta:2004}.

An alternative to monolithic HALs are the modularized or component-based HALs.
An example of component-based HAL can be found on \textsc{eCos} OS \cite{Massa:2002}.
% XXX TODO referencia melhor (mais atual) para eCos
Although, the HAL used by \textsc{eCos} is based on software components, it is
not generated according to application needs, thus it can carry unnecessary code
for the system.
In the case of SOCs generated from the micro-architecture LEON2, for example,
the \textsc{eCos} system assumes the existence of an UART device what is not
necessarily true \cite{Polpeta:2004}.
% 
As \textsc{eCos}, the AUTOSAR initiative \cite{AUTOSAR-site} proposes an approach based on a
modularized HAL.
However, some AUTOSAR implementations solve the problem of using HALs with
unnecessary code during the AUTOSAR generation process.
This is the case of the KESO JVM, which provides Java for OSEK/AUTOSAR systems \cite{Thomm:2010}.
During the configuration step of KESO, prior to linking, it is possible to select and
combine basic software modules to create an application-oriented HAL.

The development of components for abstracting hardware devices can be guided by
a methodology based on domain engineering.
That is the case of the Application-Driven Embedded System Design (ADESD)
methodology \cite{Froehlich:2001} which has been used
for abstracting hardware devices without
generating unnecessary interdependencies between the devices been abstracted.
The Embedded Parallel Operating System (EPOS) is ADESD main case study.
At EPOS the abstractions, obtained from domain decomposition, are OS abstractions
such as threads, semaphores, abstractions for communication such as network,
channels, and others.
In order to provide system abstractions with the needed hardware access, ADESD
defines the concept of \emph{hardware mediator}.
Hardware mediators sustain an interface contract between system abstractions and
the machine, allowing for such abstractions to be machine-independent \cite{Polpeta:2004}.
% 
% Another interesting alternative to the monolithic HALs approach is the concept
% of \emph{hardware mediator}, proposed by the Application-Driven Embedded System
% Design (ADESD) methodology \cite{Froehlich:2001}.
% ADESD uses domain engineering for to identify abstractions which, in the
% operating systems domain, are threads, semaphores, abstractions for communication
% such as network, channels, and others. The \emph{hardware mediator} then,
% sustain an \emph{interface contract} between system abstractions and the machine,
% allowing for such abstractions to be machine-independent \cite{Polpeta:2004}.
%

Figure \ref{fig:hardware_mediator-gs} shows hardware mediators, and their
relation to system abstractions.
Besides been componentized and application-driven, hardware mediators are
designed to be efficient.
By using metaprogramming techniques and function inlining, it is possible to
dissolve mediators in the abstractions that use them.
Thus, the overhead of HAL service invocation is avoided.

\figtwocolumn{.7}{hardware_mediator-gs}{Hardware Mediators.}

As an example, Figure \ref{progcpp:hardware_mediator_code_example} shows part of the
implementation of the \emph{CPU} hardware mediator for the IA32 architecture.
The atomic methods
\emph{finc} (\emph{Fetch and Increment}) and
\emph{fdec} (\emph{Fetch and Decrement}) are used to implement semaphore
operations \emph{v} and \emph{p}.
Both methods are implemented using inline assembly and in a header (.h) file.
Such methods are dissolved in the binding code generated by our tool,
eliminating method invocation overhead and improving performance.

\progcpp{hardware_mediator_code_example}{Implementation of the CPU mediator.}

% \subsection{FFIs}
\subsection{Hardware abstraction and MPLs}
Automatic memory management is one of the main features provided by MPLs.
Allocated objects are automatically freed after they are no longer needed,
releasing developers from writing code to deallocate objects and eliminating the
risk of memory leak.
On the other hand, because MPLs use automatic memory management, the object
address is only known by the runtime support system of the MPLs (e.g. Virtual Machine).
Therefore, MPLs are unable to directly control memory mapped devices in the
way languages such as C and C++ do (by using the concept of \emph{pointer}).
MPLs also do not provide the concept of \emph{inline assembly}, used to
control hardware devices by using dedicated I/O assembly instructions.
In other to solve these limitations, MPLs use a mechanism called Foreign
Function Interface (FFI).

FFI is a mechanism that allows for programs written in one language to use
constructs of a program written in another.
The language that defines the FFI is called \emph{host language} and the
language that has its constructs used is called \emph{guest language}.
The host language is usually a higher level language (i.e. MPL) and the guest
language is usually a lower level language, such as C and C++.
% 
The FFI mechanism has a high importance to MPLs, since runtime libraries of the
MPL are often implemented by using FFI.
This is the case of the Java Standard Platform (JSE), which uses FFI to
implement packages such as \emph{java.io}, \emph{java.net} and \emph{java.awt}.
% TODO: referencia sobre os pacotes Java e FFI...
% 
% However a FFI by itself does not guide the developers on the abstraction of
% hardware devices, it just provides for the developers a way to access
% constructions of other languages capable of access hardware devices directly.

% NOTA: de repente trocar o exemplo do sum por algo mais a ver com hardware devices...
The manual utilization of a FFI, however, is error prone.
First, developers
must take into consideration the semantics of two distinct languages
(the host and the guest language).
Second, many FFIs
% such as KNI, NanoVM FFI and Lua FFI (cite...) demands
demand
the developer to make the ``\emph{parsing}'' of each native method manually.
The following code illustrates the implementation of a sum method using the
Java FFI KNI
(a FFI used as based for many Java Virtual Machines - JVMs, targeting embedded systems)
\cite{_k_2002}.
% 
\begin{verbatim}
KNIEXPORT KNI_RETURNTYPE_VOID
Java_simplemath_Adder_sum() {
    jint a = KNI_GetParameterAsInt(1);
    jint b = KNI_GetParameterAsInt(2);
    KNI_ReturnInt(a + b);
}
\end{verbatim}
% 
In order to obtain the sum arguments, the developer must remember the order
of the formal parameter declaration, represented by the indexes ``1'' and ``2''
passed to FFI functions.

% In order to solve the problem of manual parsing of native methods arguments and
% to deal with the semantic differences between the languages handled by the FFI,
% it is possible to generate the binding code or part of it.
% 
% The task of writing binding code can be facilitate in two ways: by having an
% high level API to create bindings, and by using tools to automatically
% generating them (partially or completely).
% 
% Binding code generators solve the problem of manual parsing of native methods
% arguments and help to overcome the semantic differences between the languages
% handled by the FFI.
\emph{Binding code generators} solve the problem of manual parsing native methods
arguments and deal with the semantic differences between the languages
encompassed by an FFI.
They generate the binding code (or part of it) from the
native source code or from higher-level descriptions.
% 
The \textit{SWIG} \cite{swig-site} tool and the
Python's foreign function library \emph{ctypeslib} \cite{ctypeslib-site},
for example, generate binding code from C/C++ header files,
avoiding the manual parsing of arguments.
The former supports several languages as output, including Python, D,
and Java, and the later focuses on Python programs.
The \emph{Jeannie} language encompass the semantics of C and \java~therefore, it
is capable of perform some semantic verifications such as to check that Java members are in
fact accessible, i.e., that references to fields, methods and members types obey
their visibility (private, protected, public, or default) \cite{1297030}.
A \emph{Jeannie} program then is used to automatically generate JNI wrappers.
% Ravit et al. present a tool which aims to utilize high level features of the
% host language to be used in the binding code. Their tool generates Python
% bindings taken as input C code and interface descriptions containing,
% for example, information about functions and their parameters \cite{Ravitch:2009:AGL:1542476.1542516}.

\subsection{Discussion} \label{sota_disc}
Several FFI and binding code generator limitations had motivated us on the
elaboration of a method to abstract hardware devices for MPLs.
First, an FFI by itself does not guide the developers on the abstraction of
hardware devices, it just provides for the developers a way to access
constructs of other languages, such as \emph{pointers} and \emph{inline assembly},
capable of access hardware devices directly.
Binding code generators make easy the use of FFIs, solving the problem of manual
parsing of native methods arguments and dealing with the semantic differences
between host and guest language.
However, as FFIs, binding code generators do not deal with the problem of
device abstraction.

% XXXTODO Meio que repeteco de coisas da introdução.
% Por enquanto deixar assim, mas depois revisar.
The method presented in this paper uses the concept of hardware mediators to
abstract hardware devices and a set of FFI focused on embedded systems to
perform the interaction between hardware devices and MPLs.
The binding code is generated from descriptions of hardware mediators and
from descriptions of FFIs.
Furthermore, the adaptation of a hardware mediator to
a FFI is automatically performed, through an aspect \emph{weaving} process.
% Mencionar / Citar AOP?

A work close related to ours is presented by Schoeberl et al. \cite{Schoeberl:2011}.
By using the concept of \emph{hardware object}, it is possible to access memory
addresses from the real machine, allowing the implementation of device drivers
in Java.
This is performed by extending a special Java class (the \emph{HardwareObject} class)
which has special treatment from the point of view of the JVM.
A modified version of the JVM has special implementations for
\emph{getfield} and \emph{putfield} bytecodes, capable of accessing hardware
addresses outside the Java heap.
% However, hardware objects are platform-dependent, as in oppose to
% hardware mediators used in our approach.
% Device drivers in Java are benefited from the Java type safe and runtime checks (e.g. array bound checking)
However, the work of Schoeberl et al. focus only on Java, while our work focus on
the reuse of the functional part of a binding code (i.e. the device driver) though
several FFIs and MPLs, as explained in Section \ref{sec:proposal}.

% ------------------------------------------------------------------------------