### 2014-01-26 Papers_SVN/papers/rejected/2012/scp2012-mpl_binding_generation-mateus/tex/uart_app.tex

% ------------------------------------------------------------------------------
% NOTA TODO: Revisar / Reescrever esta subseção: Serial Communication
\subsection{Serial Communication}
% NOTA:  Acho que é interessante citar os meus papers anteriores IADIS AC 2011.
% Algo como, resultados iniciais mostrados para KESO em \cite...
The first case study evaluated as a synthetic application which uses the
\emph{Universal Asynchronous Receiver Transmitter} (UART) hardware device for
serial communication.
Such application, was evaluated in \java, using the KESO FFI, for the
architectures IA32, AVR8, and PPC32.
The Figure \ref{progjava:uart_app} shows the \java~source of the application,
which uses the UART hardware mediator in order to write characters on a serial
device.

\progjava{uart_app}{UART application.}

The \java~ class \emph{UART}, generated as one of the outputs of EBG, is the
\java~counterpart of the UART hardware mediator and has only native methods
without implementation.
The other output of EBG is the native code adapter already tailored to be
integrated with the KESO JVM.
This integration process is show in details in our previous work
\cite{Ludwich:IADIS_AC:2011}.
At such work the UART application using our approach is compared to the a
equivalent application written using the \emph{Java Standard Edition} platform,
which uses the \emph{Java Native Interface} (JNI) as FFI, running on Linux.
Our approach is around 38X faster than using JNI and Linux.

% ------------------------------------------------------------------------------