### 2014-01-26 Papers_SVN/papers/rejected/2012/todaes-hw_aspects-tiago/code_reusability_evaluation.tex

\subsubsection{Reusability evaluation}

\tab[ht]{tab_results_code_scenarios}
{Number of lines of code used to implement each scenario}

\tabTC[ht]{tab_results_code_modules}
{Number of lines of code used to implement the case studies modules}

\tabTC[ht]{tab_results_code_cs}
{Number of lines of code used to implement the scenarios adapters compared with the hand-coded
versions}

In order to provide an accurate evaluation, we have considered different code sections in the code
analysis. \emph{Declarations} are the lines of code used for class's attributes declaration, which
includes the declaration of constants, ports, modules, and signals. \emph{Behavior} includes only
the body of methods implementing behavior. \emph{Interconnection} are the lines of code used to
define the interconnection among modules and SystemC processes declaration (i.e. the content of the
constructor method of classes which inherit from \emph{sc\_module}). \emph{Other} includes the
remaining lines. Tables \ref{tab_results_code_scenarios} and \ref{tab_results_code_modules} show the
number of lines of code for each scenario and for each SystemC module, respectively. In
table \ref{tab_results_code_modules}, for modules which are implemented using more than one C++
class, the respective column includes the information for the entire class hierarchy. The number of
lines in common superclasses such as \emph{sc\_module} and \emph{Aspect\_Common} is not considered.

Table \ref{tab_results_code_cs} compares the number of lines of code of the scenario
adapters with the hand-coded versions of each case study. Column \emph{Scenarios-adapted}
shows the number of lines in the class which implements the scenario adapter. Column
\emph{Hand-coded} shows the number of lines when our approach is not used and the aspects are
hand-coded in
the core components. This column shows only the number of lines that were modified or added in
order to include the additional behavior. The results show that the effort to write a scenario
adapter is considerably smaller than directly modifying the components. The only code section that
requires more effort is not related to the actual functionality of the design, but to the syntactic
structure of SystemC (the \emph{Others} line in tables
\ref{tab_results_code_scenarios}--\ref{tab_results_code_cs}). However, even when considering all
code sections, the difference between the scenario-adapted and the hand-coded case studies ranges
from $33\%$ to $71\%$.

Table \ref{tab_results_code_total} shows the total number of lines of code when the three case
studies are considered together. These results show that the \emph{total} number of lines written
is actually smaller in the handed-coded design, highlighting the initial development overhead of
providing a modular implementation of aspects and scenarios. This reusability is quantified by
column \emph{Reusability factor} which shows that $33\%$ of the written code is reused in at least
two of our three case studies. This factor is null in the hand-coded implementations.

\tab[ht]{tab_results_code_total}
{Total number of lines of code and reusability factor }