### 2014-01-26 Papers_SVN/papers/submitted/micpro-aspects_systemc-tiago/elsevier_submission/tab_results_fpga_sim.tex

\begin{tabular}{ccccccccc}
\toprule

\multirow{2}{*}{Case study} & \multicolumn{2}{c}{Scenario-adapted} & \multicolumn{2}{c}{Hand-coded} & \multicolumn{2}{c}{Difference}\\
\cmidrule(l){2-3}
\cmidrule(l){4-5}
\cmidrule(l){6-7}
           & Slices & LPD(ns)               & Slices & LPD(ns)            & Slices   & LPD \\
\midrule
Case 1     & 651    & 10.30                 & 645    & 10.42              & 0.93\%   & -1.15\% \\
Case 2     & 133    & 07.93                 & 130    & 07.70              & 2.30\%   & 2.98\%  \\
Case 3     & 139    & 11.46                  & 135    & 11.40             & 2.96\%   & 0.52\%  \\
\bottomrule
\end{tabular}