### 2014-01-26 Papers_SVN/papers/submitted/tc-unified_design-tiago/paper_appendix_CPP.tex

\section{Review of C++ template metaprogramming}

C++ is bla bla bla.

This appendix provides a summary of C++ template metaprogramming.  The reader may also refer to
\cite{Czarnecki:2000} and \cite{Andrei:2001} for an indepth explanation.

\subsection{C++ OOP features}

\subsection{C++ templates}

What is a template ? Show and explain a function/method template.

\subsubsection{Class templates}

Show class template

\subsubsection{Template specialization}

Show two examples: partial and full specialization

\subsection{Static metaprogramming}

What the hell is this ?

\subsubsection{Examples}

\noindent\textbf{Factorial: }

\noindent\textbf{Conditional: }

\noindent\textbf{Traits:}