### 2014-01-26 Papers_SVN/papers/submitted/tc-unified_design-tiago/tab/tab_result_hw_others_area.tex

\begin{tabular}{@{}l@{\hspace{1mm}}c@{\hspace{3mm}}c@{\hspace{3mm}}c@{\hspace{3mm}}c@{\hspace{3mm}}c@{\hspace{3mm}}c@{}}
\toprule
                    & Switch & Tone gen. & Eth MAC & CPU node & RTSNoC \\
\midrule
6-input LUT         & ~~93   & ~519      & 3424    & 8312     & 2256    \\
Flip-flops          & ~108   & ~544      & 4967    & 7243     & ~689    \\
36Kb BRAM           & ~~~0   & ~~~0      & ~~~0    & ~~~2     & ~~~0    \\
\midrule
Avg. rsc. util.     & 0.03   & 0.13      & 1.11    & 2.01     & 0.30    \\
\bottomrule
\end{tabular}